---
title: |
       | 09 Estimating Cost Functions
author: "Fabian Blasch"
date: "`r format(Sys.Date(), format = '%m/%d/%Y')`"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{amssymb}
   - \usepackage{float}
   - \usepackage{titling}
   - \usepackage{xcolor}
output: 
   pdf_document:
      number_sections: TRUE
---

# Load Data

```{r, results = "hide", message = FALSE, warning = FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE,
                      fig.pos = "center",
                      fig.width = 8,
                      fig.height = 4,
                      fig.pos = "H")

# source AUX
source("./../Misc/Auxilliary.R")
source("./../Misc/model_eval.R")

# packages
get.package(c("lubridate", "glmnet", "glmnetUtils", "tidyverse", "EnvStats",
              "kdensity"))

# load data
dat_bids <- readRDS("./../../Data/Bid Tab RDS/Bids_df_split.RDS")
dat_aucs <- readRDS("./../../Data/Bid Tab RDS/Aucs_df_split.RDS")
dat_bids_ind <- readRDS("./../../Data/Bid Tab RDS/Bids_id_df.RDS")

# bid amount counted in thousands
dat_bids_ind[, "Total_Bid"] <- dat_bids_ind[, "Total_Bid"] / 1000
```

# ECDFS

```{r, fig.height = 12, fig.pos = "center"}
# split bids by competing firms and then again by auction
split(dat_bids_ind, dat_bids_ind[, "N_Comp"]) -> split_ncomp

# align
par(mfrow = c(5, 2))

# boxplots
invis.Map(\(x, n){
  
  # ecdf
  df <- ecdf(x[, "Total_Bid"])
  
  # plot
  plot(df, main = paste0("Competitors = ", n), col = 4) 
  
  # ret
  return(df)
  
}, split_ncomp, 0:(length(split_ncomp) - 1)) -> bid_n_ecdfs
```

# Estimation

First we define the expected profit from procurement as

\[
 \pi(b, x) = (b - x)H(b)
\]

where b is the bid and x is the signal associated with the contract at hand.

Let $H(b)$ be the cumulative distribution function of the lowest bid, the derivation from the cdf of the bids is trivial.

\[
  H(b) = 1 - (1 - F_b(b))^{n}
\]

and thus 

\[
  h(b) = \frac{\partial H(b)}{\partial b} =  n(1 - F(b))^{n-1}f(b)
\]

where $n$ represents the number of competitors within an auction. Estimates are obtained for $n \in [0, 9] \cap \mathbb{N}$. Hence, $H(b)$ and $h(b)$ are directly identified from the data.

Accordingly we need to estimate the density as well as the cdf of bids from the data, conditional on auction characteristics and the nuber of competitors $n$. As we observe the bids, we may use a kernel regression to first estimate the density of the bids to then derive the cdf from the density. Subsequently, we may then obtain $H(b)$ and $h(b)$.

## Densities via Base R and EnvStats

```{r, fig.height = 12}

## Estimation using base R density (implemented via EnvStats)

# align
par(mfrow = c(5, 2), mar = c(5, 4, 4, 2) + 0.1)

# Bid density functions
invis.Map(\(inp, n){

  # ecdf
  dens <- \(x) demp(x, inp[, "Total_Bid"])
  
  # plot
  hist(inp[, "Total_Bid"], main = paste0("Competitors = ", n), col = 4, ylab = "",
       prob = TRUE, breaks = length(inp[, "Total_Bid"]) / 2, 
       xlab = "")
  
  # lines
  # lines(dens(seq(1e5, 20e6, 1000)), col = "cornflowerblue") 
  lines(density(inp[, "Total_Bid"]), col = "forestgreen", lwd = 2) 
  
  # ret
  return(dens)
  
}, split_ncomp, 0:(length(split_ncomp) - 1)) -> bid_n_dens_1

```

## Densities via kdensity

```{r, fig.height = 12}
# align
par(mfrow = c(5, 2), mar = c(5, 4, 4, 2) + 0.1)

# Bid density functions
invis.Map(\(inp, n){

  # ecdf
  dens <- kdensity(inp[, "Total_Bid"], start = "exponential", kernel = "triweight", 
                   support = c(0, Inf))
  
  # plot
  hist(inp[, "Total_Bid"], main = paste0("Competitors = ", n), col = 4, ylab = "",
       prob = TRUE, breaks = length(inp[, "Total_Bid"]) / 2, 
       xlab = "")
  
  # lines
  lines(dens, col = "forestgreen", lwd = 2) 
  
  # ret
  return(dens)
  
}, split_ncomp, 0:(length(split_ncomp) - 1)) -> bid_n_dens_2
```

## Obtain CDFs from densitiy function

```{r}
invis.Map(\(inp, n){
  
  # integrate to obtain cdf
  cdf <- \(x) integrate(inp(x), lower = 0, upper = x)$value 
  
}, bid_n_dens_2, 0:(length(bid_n_dens_2) - 1)) -> emp_cdfs
```




