---
title: "First-Price Procurement Auctions"
output: 
  github_document:
    pandoc_args: --webtex
    toc: False
    toc_depth: 2
    number_sections: False
---

```{r, results = "hide", message = FALSE, warning = FALSE, include = FALSE}
# knitr options
knitr::opts_chunk$set(echo = TRUE,
                      fig.pos = "center",
                      fig.width = 8,
                      fig.height = 6,
                      fig.pos = "H")
```

This repository contains my master thesis on first-price sealed-bid procurement 
auctions. In particular, the aim is to predict award prices of auctions held by the [Colorado Department of Transportation](https://codot.gov/business/bidding/bid-tab-archives). To date, the following models with varying pre-processing schedules have been compared:

* Elastic Net
* Random Forest
  - [including logistic PCA preprocessing](https://github.com/Base-R-Best-R/Auction/blob/main/Code/Models/Colab/Nested_CV_PCA_RF.ipynb)
* [XGBoost](https://github.com/Base-R-Best-R/Auction/blob/main/Code/Models/Colab/XGboost.ipynb)

# Best Model (07/04/2022)

```{r, echo = FALSE}
# data
dat_per_plot <- readRDS("./Data/Misc Data/Figure_Data/Lasso_Performance.RDS")

# boxplot
boxplot(Award_Price ~ Model, xlab = "", ylab = "Award Price",
        ylim = c(0, 20e3), pch = 19, outcol = "darkorchid", 
        col = c("cornflowerblue", "forestgreen", "firebrick"), main = "Quality of Prediction",
        data = dat_per_plot)
```


# Required Software

## Installing Tabulizer for R > 4.1 
Unfortunately, the implementation of the Java library tabula which is a package called tabulizer cannot be installed via *install.packages()* and further the installation [guide](https://github.com/ropensci/tabulizer) is not up to date. However, with a small adaption the installation works almost as described in the installation guide for windows 10.

```{r, echo = FALSE}
sessionInfo()
```

Three steps have to be altered:

* When using Chocolately to install Java via the command prompt specify `choco install jdk8 -y` instead of `choco install jdk7 -y` 
* Within R change `Sys.setenv(JAVA_HOME = "C:/Program Files/Java/jdk1.8.0_92")` to `Sys.setenv(JAVA_HOME = "C:/Program Files/Java/jdk1.8.0_211")`
* Then install via `remotes::install_github(c("ropensci/tabulizerjars", "ropensci/tabulizer"), INSTALL_opts = "--no-multiarch")`, **after** installing *rJava* the usual way, i.e., via `install.packages()`